### 2020-04-22 final_report/methodology.tex
harmonize references in favour of \cref

In our experiments, we evaluate and compare surrogates in effort to
optimise against metrics described in \cref{sec:metrics}. To attain meaningful
and practically usable results, we require a sufficiently large and diverse pool
of surrogate classes to draw from. This is described by the listing in
\cref{tbl:surrogates}. The presented selection of models includes basic
techniques suitable for linear regression enhanced by the kernel trick or dimension
lifting, methods driven by decision trees, instance-based learning models,
ensemble regressors, randomized algorithms, neural networks and mathematical approaches
developed specifically for the purposes of surrogate modelling. For each of
these classes, a state-of-the-art implementation was selected and adapted to
operate with TBR samples.

@@

In some rows of the table, a single class in reality represents a multitude of
fundamentally similar approaches. Good examples of this are kriging methods and neural
networks. In the case of the former, multiple algorithms such as kriging based on
partial least squares and gradient-enhanced kriging are considered. In the
latter, a host of parametric graphs are defined to realise simplistic network
architecture search (see \cref{fig:nn-archs} for details). Discrimination
between such options is considered an additional hyperparameter of the
corresponding surrogate class.

>>>>

In our experiments, we evaluate and compare surrogates in effort to
optimise against metrics described in~\cref{sec:metrics}. To attain meaningful
and practically usable results, we require a sufficiently large and diverse pool
of surrogate classes to draw from. This is described by the listing in~\cref{tbl:surrogates}. The presented selection of models includes basic
techniques suitable for linear regression enhanced by the kernel trick or dimension
lifting, methods driven by decision trees, instance-based learning models,
ensemble regressors, randomized algorithms, neural networks and mathematical approaches
developed specifically for the purposes of surrogate modelling. For each of
these classes, a state-of-the-art implementation was selected and adapted to
operate with TBR samples.

@@

In some rows of the table, a single class in reality represents a multitude of
fundamentally similar approaches. Good examples of this are kriging methods and neural
networks. In the case of the former, multiple algorithms such as kriging based on
partial least squares and gradient-enhanced kriging are considered. In the
latter, a host of parametric graphs are defined to realise simplistic network
architecture search (see~\cref{fig:nn-archs} for details). Discrimination
between such options is considered an additional hyperparameter of the
corresponding surrogate class.

### 2020-04-22 final_report/results.tex
harmonize references in favour of \cref

plotted in Figure \ref{fig:sintoy} for $n=1$ and two continuous parameters $C$. ANNs trained on this model demonstrated similar performance to those on the expensive MC model. QASS performance was verified by training a 1h3f-256 ANN on the sinusoidal theory for varied quantities of initial, incremental, and MCMC candidate samples. Although the scope of this project did not include thorough searches of this hyperparameter domain, sufficient runs were made to identify some likely trends.

An increase in MCMC candidate samples was seen to have a positive but very weak effect on final surrogate precision, suggesting that the runtime of MCMC on each iteration can be limited for increased efficiency. -- Awaiting test results on initial sample quantity --. The most complex dynamics arose with the adjustment of sample increment, shown in Figure \ref{fig:qassincr}. For each tested initial sample quantity N, the optimal number of step samples was seen to be well-approximated by $\sqrt{N}$; the plotted error trends suggest that incremental samples larger than this optimum give slower model improvement on both the training and evaluation sets, and a larger minimum error on the evaluation set. This performance distinction is predicted to be even more significant when trained on the expensive MC model, where the number of sample evaluations will serve as the primary bottleneck for computation time.
\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=1.1\linewidth]{fig6a_qassincrsamp.pdf}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=1.1\linewidth]{fig6b_qassincrtime.pdf}
    \end{subfigure}
    \caption{QASS absolute training error over total sample quantity (left) and number of iterations (right). MAE represents surrogate error on the adaptively-sampled training/test set, and E\_MAE on the independent evaluation sets.}
    \label{fig:qassincr}
\end{figure}

The plateau effect in surrogate error on the evaluation set, seen in Figure \ref{fig:qassincr}, was universal to all configurations and thought to warrant further investigation. At first this was suspected to be a residual effect of retraining the same ANN instance without adjustment to data normalisation; a "Goldilocks scheme" for checking normalisation drift was implemented and tested, but did not affect QASS performance. Schemes in which the ANN is periodically retrained were also discarded, as the retention of network weights from one iteration to the next was demonstrated to greatly benefit QASS efficiency. Further insight came from direct comparison between QASS and a baseline scheme with uniformly random incremental samples, shown in Figure \ref{fig:qasssampling}.

@@

Such tests revealed that while QASS has unmatched performance on its own adaptively-sampled training set, it is outperformed by the baseline scheme on uniformly random evaluation sets. We suspected that while QASS excels in learning the most strongly peaked regions of the TBR theory, this comes at the expense of precision in broader, smoother regions where uniformly random sampling suffices. Therefore a mixed scheme was implemented, with half MCMC samples and half uniformly random samples incremented on each iteration, which is also shown in Figure \ref{fig:qasssampling}.

>>>>

plotted in~\cref{fig:sintoy} for $n=1$ and two continuous parameters $C$. ANNs trained on this model demonstrated similar performance to those on the expensive MC model. QASS performance was verified by training a 1h3f-256 ANN on the sinusoidal theory for varied quantities of initial, incremental, and MCMC candidate samples. Although the scope of this project did not include thorough searches of this hyperparameter domain, sufficient runs were made to identify some likely trends.

An increase in MCMC candidate samples was seen to have a positive but very weak
effect on final surrogate precision, suggesting that the runtime of MCMC on each
iteration can be limited for increased efficiency. -- Awaiting test results on
initial sample quantity --. The most complex dynamics arose with the adjustment
of sample increment, shown in~\cref{fig:qassincr}. For each tested initial sample quantity N, the optimal number of step samples was seen to be well-approximated by $\sqrt{N}$; the plotted error trends suggest that incremental samples larger than this optimum give slower model improvement on both the training and evaluation sets, and a larger minimum error on the evaluation set. This performance distinction is predicted to be even more significant when trained on the expensive MC model, where the number of sample evaluations will serve as the primary bottleneck for computation time.
\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=1.1\linewidth]{fig6a_qassincrsamp.pdf}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=1.1\linewidth]{fig6b_qassincrtime.pdf}
    \end{subfigure}
    \caption{QASS absolute training error over total sample quantity (left) and number of iterations (right). MAE represents surrogate error on the adaptively-sampled training/test set, and E\_MAE on the independent evaluation sets.}
    \label{fig:qassincr}
\end{figure}

The plateau effect in surrogate error on the evaluation set, seen
in~\cref{fig:qassincr}, was universal to all configurations and thought to
warrant further investigation. At first this was suspected to be a residual
effect of retraining the same ANN instance without adjustment to data
normalisation; a "Goldilocks scheme" for checking normalisation drift was
implemented and tested, but did not affect QASS performance. Schemes in which
the ANN is periodically retrained were also discarded, as the retention of
network weights from one iteration to the next was demonstrated to greatly
benefit QASS efficiency. Further insight came from direct comparison between
QASS and a baseline scheme with uniformly random incremental samples, shown
in~\cref{fig:qasssampling}.

@@

Such tests revealed that while QASS has unmatched performance on its own
adaptively-sampled training set, it is outperformed by the baseline scheme on
uniformly random evaluation sets. We suspected that while QASS excels in
learning the most strongly peaked regions of the TBR theory, this comes at the
expense of precision in broader, smoother regions where uniformly random
sampling suffices. Therefore a mixed scheme was implemented, with half MCMC
samples and half uniformly random samples incremented on each iteration, which
is also shown in~\cref{fig:qasssampling}.
